Seagrass mapping and monitoring along the coasts of Crete, Greece

### 2011-01-20 MScDiplom_Lemenkova.tex
added flowchart, p.9

\textbf{Software and utilities used in the current work}:\\
\begin{itemize}
	\item \href{http://www.esri.com/software/arcgis/index.html}{ArcGIS 10.0}: mapping, data integration and spatial analysis
	\item \href{http://www.erdas.com/products/ERDASIMAGINE/ERDASIMAGINE/Details.aspx}{Erdas Imagine}: imagery processing and classification
	\item \href{http://www.filewatcher.com/b/ftp/ftp.dfd.dlr.de/pub/WASI.0.0.html}{WASI Water Color Simulator}, for modelling optical water properties under changing environment
	\item \href{http://welcome.hp.com/country/us/en/prodserv/handheld.html}{iPAQ} and GPS for data capture during the fieldwork
	\item \href{http://www.bibtex.org/de/}{BibTeX}, for management of bibliography and reference database
	\item \href{http://www.latex-project.org/}{\LaTeX}, for document typesetting, formatting and performing pdf output
	\item \href{http://www.python.org/}{Python language}, for writing auxiliary scripts (e.g. raw data interpolation)
	\item \href{http://www.gnuplot.info/}{Gnuplot} for plotting graphs
	\item \href{http://www.openoffice.org/}{Open Office} : document typesetting, data preprocessing and preliminary statistical evalution
	\item \href{http://calc2latex.sourceforge.net/}{Calc2LaTeX} : converting spreadsheets into \LaTeX environment
	\item \href{http://www.spss.com/}{SPSS (PASW Statistics 18)} : statistical data edition and analysis 
	\item \href{http://gretl.sourceforge.net/}{Gretl: Gnu Regression, Econometrics and Time-series Library} for statistic analysis
	\item \href{http://staruml.sourceforge.net/en/}{StarUML} , \href{http://live.gnome.org/Dia}{Dia}: modelling flowchart diagrams for work algorithms
\end{itemize}
\pagebreak

@@

@@

\subsection{Research approach}
Seagrass consistent monitoring and mappint is necessary and important for the sustainable coastal
development and conservation measures. Earlier, many seagrass meadows have been destroyed by
human activities in the coastal zone, mainly due to the ignorance of their existence, because
information on the seagrass bed exact location was not available \cite{Choo06}\label{Choo06}. Well-time seagrass
observations and mapping enables precise control of its spatial distribution, detection of any changes
in the seagrass landscapes, highlightes potential environmental threats in the coastal zone (e.g.
declining of meadows) before they become unmanageable for the coastal management services.
Choosing the right and most effective approach method for the seagrass monitoring is essential.
Remote sensing methods alone, though having evident advantages, are insufficient, because satellite
images of underwater habitats are notoriously difficult to identify and interpret. The best research
method should be based on the integrated approach, well described in various scientific works \cite{Brown02,Montefalcone06,Kirkman96}\label{Brown02} \label{Montefalcone06} \label{Kirkman96}, which includes combination of various
techniques of the seagrass monitoring, i.e. remote sensing imagery classification of aerial and satellite
images, GIS-based spatial analysis and ground in-situ surveys.
The current study is based on the application of the remote sensing data, broadband satellite imagery,
aerial images and the results of the underwater videographic measurements towards seagrass
mapping. Image classification is based on the principle of the differentiation between the spectral
signatures of various seafloor cover types. The spectrum of light coming up from the ocean surface in
shallow waters keeps information on the optical properties of the seawater components and benthic
substrate which can be read from their spectral signatures \cite{Werdell03}\label{Werdell03}. Spectral
reflectances, distinct for each seafloor cover type, enable their discrimination on the image. The preprocessing
of the images includes imagery corrections for atmospheric noises and effects of the water
column. Reflectance spectra of the seagrass canopy at different depths of the water-column are
analysed for the discrimination of their spectral signatures, enabling to separate various seafloor types
during classification. The results of the of imagery classification are analysed for the detection of the
dynamics in \textit{P.oceanica} seagrass distribution along the northern coasts of Crete. Aerial imagery from
Google Earth with high spatial resolution, suitable for the large-scale detailed mapping of seagrass
mattes, is used for the improvement of the accuracy of large seagrass meadows and separate mattes
within the meadows. The in situ underwater videometric measurements of the seafloor are collected
during the fieldwork in Crete, for the validation of the classification results and to determine the exact
current distribution of the \textit{P.oceanica} meadows. The image processing includes steps of the remote
sensing techniques, i.e. calibration, masking from land and cloud, atmospheric correction, sea surface
glint and depth effects correction as recommended \cite{Matarrese08}\label{Matarrese08}. During the image classification
working step the training sites for the supervised classification methods are designed, as well as its
control and trials of different classification approaches (Unsupervised, K-means or Isodata;
Supervised, Maximum Likelihood).
\pagebreak

>>>>

\textbf{Software and utilities used in the current work}:\\
\begin{itemize}
	\item \href{http://www.esri.com/software/arcgis/index.html}{ArcGIS 10.0}: mapping, data integration and spatial analysis
	\item \href{http://www.erdas.com/products/ERDASIMAGINE/ERDASIMAGINE/Details.aspx}{Erdas Imagine}: imagery processing and classification
	\item \href{http://www.filewatcher.com/b/ftp/ftp.dfd.dlr.de/pub/WASI.0.0.html}{WASI Water Color Simulator}, for modelling optical water properties under changing environment
	\item \href{http://welcome.hp.com/country/us/en/prodserv/handheld.html}{iPAQ} and GPS for data capture during the fieldwork
	\item \href{http://www.bibtex.org/de/}{BibTeX}, for management of bibliography and reference database
	\item \href{http://www.latex-project.org/}{\LaTeX}, for document typesetting, formatting and performing pdf output
	\item \href{http://www.python.org/}{Python language}, for writing auxiliary scripts (e.g. raw data interpolation)
	\item \href{http://www.gnuplot.info/}{Gnuplot} for plotting graphs
	\item \href{http://www.openoffice.org/}{Open Office} : document typesetting, data preprocessing and preliminary statistical evalution
	\item \href{http://calc2latex.sourceforge.net/}{Calc2LaTeX} : converting spreadsheets into \LaTeX environment
	\item \href{http://www.spss.com/}{SPSS (PASW Statistics 18)} : statistical data edition and analysis 
	\item \href{http://gretl.sourceforge.net/}{Gretl: Gnu Regression, Econometrics and Time-series Library} for statistic analysis
	\item \href{http://staruml.sourceforge.net/en/}{StarUML} , \href{http://live.gnome.org/Dia}{Dia} , \href{http://argouml.tigris.org/}{AgroUML}: diagramming flowcharts for work algorithms
\end{itemize}
\pagebreak

@@

\subsection{Research approach}
Seagrass consistent monitoring and mappint is necessary and important for the sustainable coastal
development and conservation measures. Earlier, many seagrass meadows have been destroyed by
human activities in the coastal zone, mainly due to the ignorance of their existence, because
information on the seagrass bed exact location was not available \cite{Choo06}\label{Choo06}. Well-time seagrass
observations and mapping enables precise control of its spatial distribution, detection of any changes
in the seagrass landscapes, highlightes potential environmental threats in the coastal zone (e.g.
declining of meadows) before they become unmanageable for the coastal management services.
Choosing the right and most effective approach method for the seagrass monitoring is essential.
Remote sensing methods alone, though having evident advantages, are insufficient, because satellite
images of underwater habitats are notoriously difficult to identify and interpret. The best research
method should be based on the integrated approach, well described in various scientific works \cite{Brown02,Montefalcone06,Kirkman96}\label{Brown02} \label{Montefalcone06} \label{Kirkman96}, which includes combination of various
techniques of the seagrass monitoring, i.e. remote sensing imagery classification of aerial and satellite
images, GIS-based spatial analysis and ground in-situ surveys.
The current study is based on the application of the remote sensing data, broadband satellite imagery,
aerial images and the results of the underwater videographic measurements towards seagrass
mapping. Image classification is based on the principle of the differentiation between the spectral
signatures of various seafloor cover types. The spectrum of light coming up from the ocean surface in
shallow waters keeps information on the optical properties of the seawater components and benthic
substrate which can be read from their spectral signatures \cite{Werdell03}\label{Werdell03}. \\

\begin{figure}[h]
	\centering
	\includegraphics[scale=0.48]{UML_ClassDiagram.jpg}
	\caption{General methodological approach for the analysis of spectral signatures. \href{http://argouml.tigris.org/}{AgroUML}}
	\label{fig:16}
\end{figure}

Spectral reflectances, distinct for each seafloor cover type, enable their discrimination on the image. The preprocessing
of the images includes imagery corrections for atmospheric noises and effects of the water
column. Reflectance spectra of the seagrass canopy at different depths of the water-column are
analysed for the discrimination of their spectral signatures, enabling to separate various seafloor types
during classification. The results of the of imagery classification are analysed for the detection of the
dynamics in \textit{P.oceanica} seagrass distribution along the northern coasts of Crete. Aerial imagery from
Google Earth with high spatial resolution, suitable for the large-scale detailed mapping of seagrass
mattes, is used for the improvement of the accuracy of large seagrass meadows and separate mattes
within the meadows. The in situ underwater videometric measurements of the seafloor are collected
during the fieldwork in Crete, for the validation of the classification results and to determine the exact
current distribution of the \textit{P.oceanica} meadows. The image processing includes steps of the remote
sensing techniques, i.e. calibration, masking from land and cloud, atmospheric correction, sea surface
glint and depth effects correction as recommended \cite{Matarrese08}\label{Matarrese08}. During the image classification
working step the training sites for the supervised classification methods are designed, as well as its
control and trials of different classification approaches (Unsupervised, K-means or Isodata;
Supervised, Maximum Likelihood).

@@
